\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{tabularx}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}

\title{Independent Reproduction Protocol for Numerical Results}
\author{Talgat Daniyarov}
\date{}

\begin{document}
\maketitle

\section{Purpose and Decision Logic}
This protocol specifies a falsifiable reproduction plan for the paper's main claims. The intent is not to ``make the plots look good,'' but to determine whether the predicted effects appear under clean, controlled simulations and disappear under deliberate violations.

\paragraph{Notation.}
Let $m_N$ denote the effective sample size defined in the paper, and $f=n/N$ the treated (or sampled) fraction.

\paragraph{Claims and decision rules.}
Table~\ref{tab:decision-rules} lists the core claims and the expected outcomes that support or reject each one.
Claim C1 is a proxy diagnostic and does not directly test Le Cam deficiency; it checks the predicted CLT-order geometry under reduced information.

\begin{table}[ht]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabularx}{\textwidth}{Y Y Y}
\toprule
Claim & Support if (expected outcomes) & Reject if (failure modes) \\
\midrule
\textbf{C1. Gaussian-shift proxy / forced pivot (CLT-order, reduced information).} & For regular equal-tailed Bayes intervals measurable w.r.t. $(\hat\tau,\hat V)$, endpoint differences from the Gaussian pivot shrink at rate $O(m_N^{-1/2})$ uniformly across several regular priors. Gaussian-shift proxy diagnostics (QQ or distance metrics) show discrepancies decaying with $m_N$. & Endpoint differences do not decay at $m_N^{-1/2}$, or are sensitive to mild prior changes. Gaussian-shift proxy shows persistent non-Gaussian features in the asymptotic range. \\
\textbf{C2. Higher-order refinement when parity holds.} & In symmetry/parity settings, calibrated methods show $O(m_N^{-1})$ coverage error with stabilized $m_N e(m_N)$ plots and slopes near $-1$. When parity fails, errors align with $O(m_N^{-1/2})$ and $m_N^{1/2}e(m_N)$ stabilizes instead. & No regime separation: calibrated methods fail to improve under parity, or show $O(m_N^{-1})$ even when theory predicts only $O(m_N^{-1/2})$. \\
\textbf{C3. Sampling-fraction skew and near-census transition.} & Estimated skew component is near zero at $f=1/2$, flips sign across $f=1/2$, and grows with $|f-1/2|$. Near census ($1-f\asymp m_N^{-1/2}$), rate-scaled plots switch so $m_N^{1/2}e(m_N)$ stabilizes while $m_N e(m_N)$ does not. & Skew component does not vanish at $f=1/2$, no sign flip occurs across $f=1/2$, or no rate transition appears near census. \\
\textbf{C4. Lattice correction.} & Jittered calibration substantially reduces lattice periodicity (e.g., CDF jump size or coverage oscillation amplitude) and improves coverage stability relative to un-jittered Cornish--Fisher. & Periodicity persists after jittering, or jittering changes coverage negligibly while periodic artifacts remain. \\
\textbf{C5. Cluster leverage regimes.} & Under vanishing maximal leverage, calibrated methods show the predicted refinement; under leverage violation, refinement degrades and rate-scaled plots fail to stabilize. & Regime B behaves like the violation case, or violation cases still show strong refinement. \\
\textbf{C6. One-sided calibration (asymmetric sets).} & One-sided calibrated intervals show improved coverage relative to uncalibrated one-sided pivots, with rate-scaled plots consistent with the predicted $O(m_N^{-1})$ behavior under parity-holds settings. & Calibration fails to improve one-sided coverage, or errors remain at $O(m_N^{-1/2})$ despite parity-holds conditions. \\
\textbf{C7. Objective Bayes missing-mass alignment.} & Under exchangeable arm-wise priors (e.g., Bayesian bootstrap/DP) with neutral coupling, posterior variance for $\tau_N$ matches the Neyman conservative variance (missing-mass factor) up to $o(m_N^{-1})$, and equal-tailed posterior intervals align with the pivot at $O(m_N^{-1/2})$. & Posterior variance deviates systematically from the missing-mass factor, or interval endpoints diverge from the pivot at CLT order. \\
\bottomrule
\end{tabularx}
\caption{Decision rules for supporting or rejecting the main claims.}
\label{tab:decision-rules}
\end{table}

\section{Reproducibility Standards and Outputs}
\subsection{Hermetic run description}
The reproducer must log, in a single run record:
\begin{itemize}[leftmargin=*]
\item operating system and CPU model,
\item RNG type (name only),
\item exact pseudo-random seed(s) (outer population, inner assignment, jitter, and posterior sampling seeds),
\item execution date/time,
\item number of parallel workers.
\end{itemize}
All randomization must be seed-controlled and logged.

\subsection{Mandatory deliverables}
\begin{enumerate}[label=(\Alph*),leftmargin=*,itemsep=2pt]
\item \textbf{Master table} (CSV-like) with one row per configuration including: design type, outcome type, $N$, $m_N$, $f=n/N$, cluster summaries (if relevant), Monte Carlo counts $(B,R,S)$, coverage and MCSE for each method (two-sided and one-sided where applicable), average interval length, posterior variance ratio (posterior variance divided by Neyman conservative variance), and diagnostics (pivot skew/kurtosis, lattice periodicity metric, maximal leverage $\lambda_N$).
\item \textbf{PDF report} with: coverage vs. $m_N$ plots, log--log slope checks, sampling-fraction and near-census panels, one-sided calibration panels, lattice periodicity before/after jitter, clustered leverage plots, Objective Bayes variance-ratio and interval-alignment panels, violation stress tests, and a run-log appendix.
\end{enumerate}

\section{Experimental Designs and DGPs}
\subsection{Designs}
\begin{enumerate}[label=(D\arabic*),leftmargin=*,itemsep=2pt]
\item CRA with treated fraction grid $f\in\{0.20,0.35,0.50,0.65,0.80,0.90\}$ (include $f=1/2$ exactly) and a near-census sequence with $1-f\asymp m_N^{-1/2}$ (e.g., $f=1-c/\sqrt{m_N}$ for $c\in\{1,2\}$).
\item Stratified design with at least two strata and heterogeneous stratum variances; include at least two $f$ values on opposite sides of $1/2$.
\item Clustered designs: Regime A (bounded cluster size) and Regime B (growing clusters with vanishing maximal leverage).
\item Violation case: at least one clustered configuration with $\lambda_N$ bounded away from $0$.
\end{enumerate}

Report the effective sample size $m_N$ as defined in the paper for each design and include its value in the master table.

\subsection{Outcome regimes}
At minimum:
\begin{itemize}[leftmargin=*]
\item \textbf{Skewed continuous outcomes:} e.g., lognormal or $\chi^2$-type finite populations.
\item \textbf{Symmetric continuous outcomes (parity-holds):} construct finite populations with near-zero standardized third cumulant (e.g., pair values symmetrically around the mean).
\item \textbf{Binary outcomes:} Bernoulli finite populations (lattice setting).
\end{itemize}
For each configuration, draw a finite population once per outer draw, then randomize assignments repeatedly in the inner loop.

\subsection{Parity modules}
Construct paired continuous-outcome configurations to separate parity-holds and parity-fails regimes:
\begin{itemize}[leftmargin=*]
\item \textbf{Parity-holds:} symmetric outcomes with $f=1/2$; verify the standardized third cumulant of the pivot (or Hájek-projected array) is near zero.
\item \textbf{Parity-fails:} skewed outcomes and/or $f\neq 1/2$; verify nonzero standardized third cumulant.
\end{itemize}
Use the same $N$ grid for both to enable rate comparisons.

\subsection{FPC sanity module}
Generate a finite population $y_{1:N}$, draw an SRSWOR sample of size $n$, and verify
\[\frac{\widehat{\mathrm{Var}}(\mu_N\mid y_S)}{(1-f)s_S^2/n}\approx 1\]
across several $(N,n)$ pairs spanning $f\in[0.2,0.8]$.

\section{Procedures to Compare}
\subsection{Baseline}
Gaussian pivot interval: $\hat\tau \pm z_{1-\alpha/2}\,\widehat{\mathrm{SE}}$.

\subsection{Higher-order calibration (continuous)}
Cornish--Fisher calibrated interval using estimated cumulants/Edgeworth terms.

\subsection{Higher-order calibration (lattice)}
Cornish--Fisher with lattice jittering: add independent uniform jitter with half-width equal to half the lattice span in standardized units. Use an independent jitter seed, logged.

\subsection{Forced-pivot diagnostic (C1)}
Construct equal-tailed Bayesian intervals based on the pivot likelihood with several regular priors on the standardized parameter (e.g., normal, Student-$t$ with moderate degrees of freedom, Laplace). Compare endpoint differences to the Gaussian pivot and verify $O(m_N^{-1/2})$ convergence uniformly across priors.

\subsection{Gaussian-shift proxy diagnostic (C1)}
Compare the empirical distribution of $(\hat\tau,\hat V)$ to the Gaussian-shift benchmark (e.g., QQ plots and a Wasserstein or Kolmogorov distance). Verify decay of the discrepancy with $m_N$.

\subsection{One-sided calibration (C6)}
Construct one-sided intervals using the explicit calibration operator for asymmetric sets. Compare coverage and rate-scaled plots to the uncalibrated one-sided pivot interval.

\subsection{Objective Bayes posterior (C7)}
Implement arm-wise exchangeable priors (e.g., Bayesian bootstrap or DP) with neutral coupling across potential outcomes. Compute the posterior for $\tau_N$ and its variance, compare the variance to the Neyman conservative variance, and compare equal-tailed posterior intervals to the Gaussian pivot at CLT order.

\section{Monte Carlo Design and MCSE Reporting}
Use a two-level simulation:
\begin{itemize}[leftmargin=*]
\item Outer loop: $B$ independent finite populations.
\item Inner loop: $R$ assignments per population.
\end{itemize}
Report MCSE for coverage using both within-population and between-population components; include the combined MCSE in the master table.

\section{Rate Verification and Diagnostics}
\subsection{Rate plots}
For each design/outcome/method, define coverage error $e(m_N)=|\hat p(m_N)-(1-\alpha)|$ and plot:
\begin{itemize}[leftmargin=*]
\item $m_N\,e(m_N)$ vs. $m_N$ (stabilization indicates $O(m_N^{-1})$),
\item $m_N^{1/2}e(m_N)$ vs. $m_N$ (stabilization indicates $O(m_N^{-1/2})$).
\end{itemize}
Estimate log--log slopes on the largest 3--4 $m_N$ points and report slope estimates with MCSE-based uncertainty.

\subsection{Sampling-fraction skew}
For CRA and one clustered configuration, run the $f$-grid and report:
\begin{itemize}[leftmargin=*]
\item estimated pivot skewness (third standardized cumulant),
\item fitted coefficient on $(1-2f)$ in the empirical skew decomposition,
\item coverage improvements as a function of $f$.
\end{itemize}
Pass criteria: the design-induced component is near zero at $f=1/2$, the sign flips across $f=1/2$, and magnitude grows with $|f-1/2|$.

\subsection{Near-census transition}
Using the near-census sequence, compare rate-scaled errors across $f=1/2$ and near-census values. Pass criteria: near census, $m_N^{1/2}e(m_N)$ stabilizes while $m_N e(m_N)$ does not, indicating the rate transition.

\subsection{Lattice periodicity}
Report a periodicity diagnostic (e.g., maximum adjacent-step jump in the pivot CDF or coverage oscillation amplitude) before and after jittering. Pass criterion: the jittered method substantially reduces periodicity.

\subsection{Cluster leverage}
Report $\lambda_N=\max_g n_g/\sum_h n_h$ across $N$ and confirm the assumed decay in Regime B. In the violation case, rate-scaled plots should fail to stabilize as predicted.

\subsection{One-sided calibration}
Report one-sided coverage errors for calibrated and uncalibrated intervals with rate-scaled plots. Pass criteria: calibrated one-sided intervals improve coverage and show the predicted scaling in parity-holds settings.

\subsection{Objective Bayes alignment}
Report the posterior variance ratio (posterior variance divided by Neyman conservative variance) and endpoint differences between Objective Bayes intervals and the Gaussian pivot. Pass criteria: variance ratios near 1 and endpoint differences shrinking at $O(m_N^{-1/2})$.

\section{Integrity Checks}
\begin{itemize}[leftmargin=*]
\item Confirm that potential outcomes are generated once per finite population (outer loop) and assignments are randomized independently (inner loop).
\item Apply lattice jittering to a continuous-outcome configuration and verify negligible effect on coverage and length.
\item Report pivot skewness and kurtosis to ensure corrections move in the predicted direction.
\item Verify parity-holds configurations have near-zero standardized third cumulant and parity-fails configurations do not.
\item For Objective Bayes runs, verify the posterior variance incorporates the missing-mass factor and matches the Neyman conservative variance to the expected order.
\item For one-sided intervals, verify calibration improves coverage relative to the uncalibrated one-sided pivot.
\end{itemize}

\section{Minimum Configuration Grid}
\begin{itemize}[leftmargin=*]
\item CRA: continuous + binary, $f\in\{0.20,0.35,0.50,0.65,0.80,0.90\}$ plus a near-census sequence with $1-f\asymp m_N^{-1/2}$, at least 5 $N$ values.
\item Parity modules: symmetric continuous outcomes at $f=1/2$ and skewed continuous outcomes at $f\neq 1/2$, at least 3 $N$ values each.
\item Stratified: continuous + binary, two $f$ values straddling $1/2$, at least 3 $N$ values.
\item Clustered Regime A and B: continuous + binary, at least 3 $N$ values, include leverage diagnostics.
\item Violation case: at least one $N$ (preferably two) with $\lambda_N$ bounded away from 0.
\item One-sided inference: at least one continuous and one binary configuration with one-sided intervals.
\item Objective Bayes: at least one CRA and one stratified configuration using exchangeable priors, report posterior variance ratios and interval alignment.
\item FPC sanity module: at least 4 $(N,n)$ pairs spanning $f\in[0.2,0.8]$.
\end{itemize}

\section{Reporting Requirements}
The final PDF report must include method definitions, coverage/length tables with MCSEs, rate-scaled plots, sampling-fraction skew and near-census transition panels, one-sided calibration panels, lattice periodicity diagnostics, cluster leverage plots (including violation case), Objective Bayes variance-ratio and interval-alignment panels, and a run-log appendix. All deviations from this protocol must be explicitly documented and justified.

\end{document}
